---
title: "Assignment5_Decomposition"
output: html_document
date: "2023-10-27"
---
## Description

The dataset set shows the Retail sales of Beer, Wine, and Liquor Stores from January 2018 to July 2023. The sales are in millions of dollars over the period mentioned above. 

The time series graph of this dataset shows the seasonality in sales every year. The sales values grow over the year, especially during the holiday season. The graph shows the spike highest during the year's end and second highest during the month mid-year i.e., May and July. The graph shows a decline in sales at the beginning of the year.

Looking at the time series data it can also be inferred that there has been a visible increasing trend.
Sales growth has been shown over the last 5 years. It has increased from 2018 to 2023.

```{r}
library(fpp)
library(fpp2)

library(readxl)
MyDf=read_excel("TimeSeriesData_LiquourSales.xls")

# Change column names to readable form
colnames(MyDf) <- c("Date", "Sales")

View(MyDf)

# Change date format from POSIXct to Date.
MyDf$Date <- as.Date(MyDf$Date)
```

## Converting Data into timeseries
```{r}
Tsales <- ts(MyDf$Sales, start = c(2018,1), end = c(2023,7), frequency = 12)
```

## Seasonal Decomposition of Time Series by Loess
```{r}
Tsales_decomp <- stl(Tsales,s.window ="periodic")
plot(Tsales_decomp)
attributes(Tsales_decomp)

```

## Plot the seasonal component

```{r}
plot(Tsales_decomp$time.series[, "seasonal"], main="Seasonal Component")
```

## Plot the trend component
```{r}
plot(Tsales_decomp$time.series[, "trend"], main="Trend Component")
```

## Plot the remainder component
```{r}
plot(Tsales_decomp$time.series[, "remainder"], main="Remainder (Residual) Component")
```

## Plot seasonal adjustment
```{r}
DecomTsale<-seasadj(Tsales_decomp)
plot(DecomTsale)

# Plot a line on the graph

plot(Tsales)
lines(DecomTsale, col="red")

```

## Default period forecast

```{r}
f_stl_forecast <- forecast(Tsales_decomp)
plot(f_stl_forecast)
```

## Forecast using period timestamp
```{r}
f_stl_timestamp <- forecast(Tsales_decomp,h=40)
plot(f_stl_timestamp)
```

## Forecasting the Seasonally Adjusted Decomposed Time Series
```{r}
DecomTsale_forecast <- forecast(DecomTsale, h=40)
plot(DecomTsale_forecast)
```
```

# Another way for Decompostion using Decompose function

```{r}
decomp_sales <- decompose(Tsales)
plot(decomp_sales)
```
# Seasonal Adjustment

```{r}
Decomp_sales_seaAJ = seasadj(decomp_sales)
plot(Decomp_sales_seaAJ)
```
## Plotting Line Seasonally Adjusted Decomposed TS with the original Time Series

```{r}
plot(Tsales)
lines(Decomp_sales_seaAJ, col="red")
```

## Forecasting the Seasonally Adjusted Decomposed Time Series

```{r}
T_sales_decompose_seasadj_forecast <- forecast(Decomp_sales_seaAJ, h=40)
plot(T_sales_decompose_seasadj_forecast)

```

## Forecast just the seasonality

```{r}
decomp_sales$seasonal
sea<-forecast(decomp_sales$seasonal)
plot(sea)
```

## Forecast random 

```{r}
decomp_sales$random
ran<-forecast(decomp_sales$random)
plot(ran)
```

## Forecast the trend

```{r}
decomp_sales$trend
trd<-forecast(decomp_sales$trend)
plot(trd)
```

#### By subtracting the seasonal component from the original time series we can obtain a seasonally adjusted time series as shown in this plot here

## Doing Seasonal Adjustment
```{r}
seasadj <- Tsales-decomp_sales$seasonal
 plot(seasadj)  
```

## Each one shows different attributes 
```{r}
attributes(decomp_sales)
seasadj(decomp_sales)
```

